---
title: "Capitol Reef Photo Data Summary"
author: "IA Student Analyst Team"
date: "2024-02-01"
output: 
  prettydoc::html_pretty:
    theme: cayman
---
```{r, include=FALSE}
# Set CRAN mirror before installing packages
options(repos = c(CRAN = "https://cloud.r-project.org"))

# Install and load the necessary packages
if (!requireNamespace("prettydoc", quietly = TRUE)) {
  install.packages("prettydoc")
}
library(prettydoc)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, include=FALSE}
library(readxl)
library(dplyr)
library(openxlsx)
library(RColorBrewer)
library(viridis)
library(ggplot2)
```

```{r, include=FALSE}
file_path <- "PhotoWildlifeData_UMichDataShare_12042023.xlsx"
sheet_names <- excel_sheets(file_path)
all_dataframes <- list()
for (sheet in sheet_names) {
  df <- read_excel(file_path, sheet = sheet)
  all_dataframes[[sheet]] <- df
}

photoData <- all_dataframes[['PhotoData']]
siteLocations <- all_dataframes[['SiteLocations']]
metaData <- all_dataframes[['Metadata']]

```
## Capitol Reef Camera Data
```{r, include= FALSE}
photoData$DateTime <- as.POSIXct(photoData$DateTime, format="%Y-%m-%d %H:%M:%S")

max_date <- max(photoData$DateTime)
min_date <- min(photoData$DateTime)

min_date <- as.Date(min_date)
max_date <- as.Date(max_date)
```
The following file is a summary of the data captured from motion sensor cameras within Capitol Reef National Park. The sightings in this report occur between `r min_date` and `r max_date`.

```{r}

```

## Including Plots

You can also embed plots, for example:

```{r pressure, include=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
