---
title: "Capitol Reef Photo Data Summary"
author: "IA Student Analyst Team"
date: "`r format(Sys.Date(), '%m/%d/%Y')`"
output: 
  prettydoc::html_pretty:
    theme: cayman
    toc: true
    toc_depth: 3
---
<!--DOWNLOAD ALL NECESSARY LIBRARIES -->
```{r, include=FALSE}
# Set CRAN mirror before installing packages
options(repos = c(CRAN = "https://cloud.r-project.org"))

# Install and load the necessary packages
if (!requireNamespace("prettydoc", quietly = TRUE)) {
  install.packages("prettydoc")
}
if (!require("kableExtra")) install.packages('kableExtra')
if (!require('plyr')) install.packages('plyr')
if (!require('tidyverse')) install.packages('tidyverse')
if (!require('plotly')) install.packages('plotly')
library(tidyverse)
library(kableExtra)
library(prettydoc)
library(dplyr)
library(ggplot2)
library(plotly)
library(readxl)
library(openxlsx)
library(RColorBrewer)
library(viridis)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#install.packages("plotly")
library(plotly)
library(tidyverse)
```

<!-- MODIFY FILE PATH HERE -->
```{r, include=FALSE}

#MODIFY FILE PATH HERE
#file_path <- "CARE_GameCam_MergedData/TimelapseData_merged.csv"
file_path <- "PhotoWildlifeData_UMichDataShare_12042023.xlsx"
file_extension <- tools::file_ext(file_path)

if(file_extension == "csv") {
  # If the file is a CSV, read it directly into photoData
  photoData <- read_csv(file_path, show_col_types = FALSE)
} else if(file_extension %in% c("xlsx", "xls")) {
  # If the file is an Excel workbook, and you specifically want the "PhotoData" sheet
  # Adjust "PhotoData" below if the sheet name is different
  if("PhotoData" %in% excel_sheets(file_path)) {
    photoData <- read_excel(file_path, sheet = "PhotoData")
  } else {
    stop("The 'PhotoData' sheet does not exist in this workbook.")
  }
} else {
  stop("Unsupported file type.")
}

metrics <- read_excel('Metrics_CameraData_capitolReefNP.xlsx')

```


## Overview: Capitol Reef Camera Data

```{r, include=FALSE}
photoData$DateTime <- as.POSIXct(photoData$DateTime, format="%Y-%m-%d %H:%M:%S")

max_date <- max(photoData$DateTime)
min_date <- min(photoData$DateTime)

min_date <- as.Date(min_date)
max_date <- as.Date(max_date)

locations <- c(unique(photoData$RegionSite))
cameras <- unique(photoData$Camera)
```

The following file is a summary of the data captured from motion sensor cameras within Capitol Reef National Park. The sightings in this report occur between `r format(min_date, '%m/%d/%Y')` and `r format(max_date, '%m/%d/%Y')`. It includes a summary from `r length(locations)` location(s) and `r length(cameras)` different camera(s).  
<br>
NOTE: Repeats are standardized by counting one detection per two-hours to account for animals that hang around the cameras.

*** 

## Bighorn Sheep Monitoring:

```{r, include=FALSE}
#IF OTHER SPECIES DESIRED MODIFY HERE
desired_species = tolower('Bighorn sheep')

#THIS FUNCTION EXCLUDES REPEATS AND GETS SPECIES1 & SPECIES2 SHEEP
filter_rows <- function(data) {
  data %>%
    mutate(Species1 = as.character(Species1),
           Species2 = as.character(Species2),
           Repeat = as.character(Repeat)) %>%
    # Filter rows where either Species1 or Species2 matches the desired_species, case-insensitively
    filter(tolower(Species1) == desired_species | tolower(Species2) == desired_species) %>%
    # Filter rows where Repeat is NA or Repeat is not 'yes', 
    filter(is.na(Repeat) | tolower(trimws(Repeat)) != "yes")
}

sheep_standardize <- filter_rows(photoData)
```

```{r, include=FALSE}
species1_rows <- sheep_standardize %>%
  select("Project", "Region", "Site", "Camera", "DateTime", "File", "Species1", "Age_Species1", "Sex_Species1", "Count_Species1", "Repeat")

species2_rows <- sheep_standardize %>%
  select("Project", "Region", "Site", "Camera", "DateTime", "File", "Species2", "Age_Species2", "Sex_Species2", "Count_Species2", "Repeat")

species2_rows <- species2_rows %>%
  rename("Species1" = Species2,
         "Age_Species1" = Age_Species2,
         "Sex_Species1"= Sex_Species2,
        "Count_Species1" = Count_Species2)

sheep_standardize <- rbind(species1_rows, species2_rows)
sheep_standardize <- sheep_standardize[order(as.Date(sheep_standardize$DateTime, format="%Y-%m-%d %H:%M:%S")), ]
```


```{r, echo = FALSE}
#SETUP STANDARDIZED SHEEP FRAME (EXCLUDED REPEATS)
# sheep_standardize$Sex_Species1 <- ifelse(sheep_standardize$Sex_Species1 %in% c('Male', 'Female'), sheep_standardize$Sex_Species1, 'Unknown')

sheep_standardize <- sheep_standardize %>%
    mutate(
    Sex_Species1 = ifelse(
      is.na(Sex_Species1) | tolower(Species1) != tolower(desired_species) | 
      !Sex_Species1 %in% c('Male', 'Female'), 
      'Unknown', 
      Sex_Species1
    )
  )

genders <- unique(c(sheep_standardize$Sex_Species1))

# sheep_standardize %>%
#   mutate(
#     across(
#       c(Count_Species1, Count_Species2),
#       ~ as.character(.)
#     )
#   ) %>%
#   rename(
#     c('Type_Species1' = 'Species1', 'Type_Species2' = 'Species2')
#   ) %>%
#   pivot_longer(
#     cols = c(Type_Species1, Type_Species2, Age_Species1, Age_Species2, 
#              Sex_Species1, Sex_Species1, Count_Species1, Count_Species2),
#     names_to = c('.value', '.set'),
#     names_sep = '_'
#   ) %>%
#   filter(!is.na(Type))

```


### Detection of BHS Through Time


```{r, echo= FALSE}
##Detection of BHS through time

sheep_standardize$Count_Species1 <- as.numeric(as.character(sheep_standardize$Count_Species1))

daily_sightings <- sheep_standardize %>%
  # Assuming 'datetime' is in POSIXct format; otherwise, convert it
  mutate(Date = as.Date(DateTime)) %>%
  group_by(Date) %>%
  summarise(Sightings = sum(Count_Species1), .groups = 'drop')

#complete_sightings$Sightings[is.na(complete_sightings$Sightings)] <- 0


# Now plot the data
p0 <- ggplot(daily_sightings, aes(x = Date, y = Sightings)) +
  geom_line(color = "seagreen4") +  # Set the line color to green
  geom_point(color = "black") + 
  labs(title = "Daily Sheep Sightings Over Time",
       x = "Date",
       y = "Number of Sightings") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) 

p0 <- ggplotly(p0)
p0
```




```{r, echo = FALSE}
gender_sightings <- sheep_standardize %>%
  group_by(DateTime, Sex_Species1) %>%
  summarise(
    Sightings = sum(Count_Species1, na.rm = TRUE),
    .groups = 'drop'
  )

# Create the facet plot
pg <- ggplot(gender_sightings, aes(x = DateTime, y = Sightings, color = Sex_Species1)) +
  geom_line() +  # Create a line plot
  geom_point() +  # Add points
  scale_color_brewer(palette = "Paired") +  # Use the 'Paired' color palette for lines
  geom_point(color = "black") +
  facet_wrap(~Sex_Species1, scales = "free_y", ncol = 1) +  # Facet by gender
  labs(title = "Sheep Sightings Over Time by Gender",
       x = "Date",
       y = "Number of Sightings") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),  # Angle the date labels for readability
        legend.position = "bottom") +
  guides(color = guide_legend(title = "Sex"))

pg <- ggplotly(pg)
pg
```

### Overall Count & Percentage of Bighorn Sheep detection by sex:

The following graph demonstrates the overall detection of Bighorn Sheep throughout all recorded locations between `r format(min_date, '%m/%d/%Y')` and `r format(max_date, '%m/%d/%Y')`.

```{r, echo = FALSE}
#OVERALL COUNT OF BHS
coul <- brewer.pal(length(genders), "Paired")

counts <- sheep_standardize %>%
  group_by(Sex_Species1) %>%
  summarise(Total = sum(Count_Species1, na.rm = TRUE)) %>%
  ungroup() 

counts <- setNames(counts$Total, counts$Sex_Species1)

# midpoints <- barplot(counts, main="Overall Count of Bighorn Sheep Detection by Sex",
#    xlab="Sex",
#    ylab= 'Total Count',
#    col=coul,
#    ylim=c(0, max(counts) + 10),
#    font.main=1
#    )
# text(midpoints, counts, labels=counts, pos=3, cex=0.8)

interactive_bar <- plot_ly(
  x = names(counts),
  y = counts,
  type = 'bar',
  marker = list(color = coul),
  name = 'Total Count'
) %>%
  layout(
    title = "Overall Count of Bighorn Sheep Detection by Sex",
    xaxis = list(title = "Sex"),
    yaxis = list(title = "Total Count"),
    barmode = 'group'
  )
interactive_bar

```

```{r, echo=FALSE}
#PIE CHART OF BHS DETECTIONS
total_counts <- sum(counts)
percentages <- round(100 * counts / total_counts, 1)

#lables with percentages
labels <- paste(names(counts), "-", percentages, "%", sep="")

# Plot pie chart
pie(counts, labels = labels, col = brewer.pal(length(counts), "Paired"),
    main="Overall Percentage of Bighorn Sheep Detection by Sex", font.main=1)



```

### Parkwide Annual Detection of Bighorn Sheep by Sex

The following has the percentage of bighorn sheep detection by gender for the years recorded in the provided files.

```{r,fig.width= 8, echo=FALSE}
#PERCENTAGE OF BHS DETECTIONS FOR YEARS RECORDED BY SEX

# aggregate counts by year and sex
sheep_standardize$Year <- as.numeric(format(as.POSIXct(sheep_standardize$DateTime, tz = "UTC"), "%Y"))
unique_years <- unique(sheep_standardize$Year)
unique_years <- sort(unique_years)

#get the yearly totals
yearly_totals <- sheep_standardize %>%
  dplyr::group_by(Year) %>%
  dplyr::summarise(total_counts = sum(Count_Species1), .groups = 'drop')

# Then, calculate counts per Sex_Species1 within each year
detailed_counts <- sheep_standardize %>%
  dplyr::group_by(Year, Sex_Species1) %>%
  dplyr::summarise(count = sum(Count_Species1), .groups = 'drop')

# Join the totals with the detailed counts to calculate percentages
plot_df <- detailed_counts %>%
  left_join(yearly_totals, by = "Year") %>%
  mutate(percent = round(100 * count / total_counts, 1))

p <- ggplot(plot_df, aes(x = factor(Year), y = percent, fill = Sex_Species1)) +
  geom_bar(stat = "identity", position = position_dodge(), width = 0.7) + # Use dodge to place bars side by side
  geom_text(aes(label = paste(percent, "%")), position = position_dodge(width = 0.6), vjust = -0.25, size=3) +
  scale_fill_brewer(palette = "Paired", name = "Gender") + # Set a color palette and name the legend as "Gender"
  labs(title = "Annual Percentage of Bighorn Sheep by Year and Gender (Parkwide)",
       x = "Year", 
       y = "Percentage") +
  theme_minimal() + # Use a minimal theme for a cleaner look
  theme(legend.title = element_text(size = 12), # Adjust legend title size
        legend.text = element_text(size = 10), # Adjust legend text size
        plot.title = element_text(hjust = -0.5)) # Center the plot title

p_interactive <- ggplotly(p, tooltip = c("percent", "Sex_Species1"))
# Display the interactive plot
p_interactive



```

### Parkwide Annual & Monthly Detection of Bighorn Sheep by Sex

The following has the percentage of bighorn sheep detection by gender specified by the year and months recorded in the provided files.

```{r,fig.width= 8, include=FALSE}
#PERCENTAGE OF BHS (YEAR AND MONTH)

# extract the month
sheep_standardize$Month <- as.numeric(format(as.POSIXct(sheep_standardize$DateTime, tz = "UTC"), "%m"))

unique_months <- unique(sheep_standardize$Month)
unique_months <- sort(unique_months)


#monthly totals
monthly_totals <- sheep_standardize %>%
  dplyr::group_by(Year, Month) %>%
  dplyr::summarize(total_counts = Count_Species1, .groups = 'drop')

# Calculate detailed counts by year and month for each sex
detailed_counts <- sheep_standardize %>%
  group_by(Year, Month, Sex_Species1) %>%
  summarise(count = Count_Species1, .groups = 'drop')

# Join the totals with the detailed counts for annual data
plot_df_annual <- detailed_counts %>%
  group_by(Year, Sex_Species1) %>%
  summarise(count = sum(count), .groups = 'drop') %>%
  left_join(yearly_totals, by = "Year") %>%
  mutate(percent = round(100 * count / total_counts, 1))

# Join the totals with the detailed counts for monthly data
plot_df_monthly <- detailed_counts %>%
  left_join(yearly_totals, by = c("Year")) %>%
  mutate(percent = round(100 * count / total_counts, 1))

```


```{r, echo=FALSE}
#This piece of code allows us to fill in the months that are missing with 0s
complete_months <- expand.grid(
  Year = unique(plot_df_monthly$Year),
  Month = factor(1:12, labels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")),
  Sex_Species1 = unique(plot_df_monthly$Sex_Species1)
) 

plot_df_monthly$Month <- factor(plot_df_monthly$Month, levels = 1:12, labels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))

plot_df_monthly_complete <- complete_months %>%
  left_join(plot_df_monthly, by = c("Year", "Month", "Sex_Species1")) %>%
  replace_na(list(count = 0, percent = NA))  # Adjust with the appropriate column names

dodge_width <- 0.7

p2 <- ggplot(plot_df_monthly_complete, aes(x = Month, y = percent, fill = Sex_Species1)) + 
  geom_bar(stat = "identity", position = position_dodge()) + 
  facet_wrap(~Year, ncol = 1, scales = "free_y") + 
  scale_fill_brewer(palette = "Paired", name = "Gender") + 
  labs(title = "Monthly Percentage of Bighorn Sheep Detections by Gender and Year",
       x = "Month", 
       y = "Percentage") + 
  theme_minimal() + 
  theme(axis.text.x = element_text( vjust = 1, hjust=1))



p2 <- ggplotly(p2) 
p2
```

### Regional Detection of Bighorn Sheep by Sex

The following has the percentage of bighorn sheep detection by gender for the regions in the provided files.

```{r, include = FALSE}

# Calculate total counts for the entire park
total_region_sightings <- sum(sheep_standardize$Count_Species1)

# Calculate counts and percent for each region and sex against the park total
sheep_gender_by_region <- sheep_standardize %>%
  group_by(Region, Sex_Species1) %>%
  summarise(count = sum(Count_Species1), .groups = 'drop') %>%
  ungroup() %>%
  mutate(proportion = round(100 * count / total_region_sightings, 1))

```

```{r, echo=FALSE}
p3 <- ggplot(sheep_gender_by_region, aes(fill = Sex_Species1, y = proportion, x = Region)) +
  geom_bar(position = "fill", stat = "identity") +
  scale_y_continuous(labels = scales::percent_format()) +
  scale_fill_brewer(palette = "Paired") +
  labs(title = "Percentage of BHS Sex Relative to Region Totals",
       x = "Region",
       y = "Percentage") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggplotly(p3)
# Display the interactive plot
```



```{r, include=FALSE}
p4 <- ggplot(sheep_gender_by_region, aes(x = Region, y = proportion, fill = Sex_Species1)) +
  geom_bar(stat = "identity") +
  scale_fill_brewer(palette = "Paired") +
  labs(title = "Percentage of Bighorn Sheep by Sex and Region Relative to Park Totals",
       x = "Region",
       y = "Percentage",
       fill = "Sex") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "bottom")

ggplotly(p4)

```
### Annual Percentage of BHS Detections by Region and Sex
```{r, echo=FALSE, warning=FALSE}
sheep_gender_by_region <- sheep_standardize %>%
  group_by(Year, Region, Sex_Species1) %>%
  summarise(count = Count_Species1, .groups = 'drop') %>%
  left_join(yearly_totals, by = c("Year")) %>%
  mutate(percent = round(100 * (count / total_counts), 1))

#Visualization for sheep totals
dodge <- position_dodge(width=0.8)
p <- ggplot(sheep_gender_by_region, aes(x = Region, y = percent, fill = Sex_Species1)) + 
  geom_bar(stat = "identity", position = dodge, width = 0.7) + 
  facet_wrap(~Year, ncol = 1, scales = "free_y") + 
  scale_fill_brewer(palette = "Paired", name = "Gender") + 
  labs(title = "Annual Percentage of BHS Detections by Gender and Region relative to Yearly Totals",
       x = "Region", 
       y = "Percentage") + 
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust = 0.5),
        legend.position = "bottom",
        panel.spacing = unit(2, "lines")) # Adjust spacing between panels

# Convert the ggplot object to a plotly object
p_interactive <- ggplotly(p, tooltip = "label+percent+total_counts")
p_interactive <- ggplotly(p, tooltip = c("Region", "percent", "Sex_Species1", "total_counts"))
p_interactive

```
### Percentage of BHS by Region, Year, Month, & Sex

The following graphs demonstrate what percentage of BHS were detected in a particular region based on the regional yearly detections. This is further broken down to the month and the sex.

```{r, echo = FALSE}
yearly_totals <- sheep_standardize %>%
  group_by(Year, Region) %>%
  summarise(TotalYearlyDetections = sum(Count_Species1), .groups = 'drop')

monthly_detections <- sheep_standardize %>%
  group_by(Year, Month, Region, Sex_Species1) %>%
  summarise(Detections = sum(Count_Species1), .groups = 'drop')


# Calculate detections per year, month, region, and gender
monthly_percentages <- monthly_detections %>%
  # group_by(Year, Month, Region, Sex_Species1) %>%
  # summarise(Detections = n(), .groups = 'drop') %>%
  left_join(yearly_totals, by = c("Year", "Region")) %>%
  mutate(Percentage = round(100 * (Detections / TotalYearlyDetections), 1))

fill_missing_months <- function(data) {
  # Define all the months. Assuming your Month column is numeric with values 1-12
  all_months <- 1:12

  # Expand the data to include all year-month-region-sex combinations
  data %>%
    # Ensure Year, Region, and Sex_Species1 are factors so that all levels are kept even if some are missing
    mutate(Year = factor(Year),
           Region = factor(Region),
           Sex_Species1 = factor(Sex_Species1),
           Month = factor(Month, levels = all_months)) %>%
    # Complete the missing combinations
    complete(Year, Region, Sex_Species1, Month, fill = list(Detections = NA, TotalDetections = NA)) %>%
    # Arrange the data for better readability
    arrange(Year, Region, Sex_Species1, Month)
}


# Use the function on your monthly_percentages data frame
monthly_complete <- fill_missing_months(monthly_percentages)

monthly_complete$Month <- factor(monthly_complete$Month, levels = 1:12, labels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))

# Create a list of unique regions



```

```{r, echo=FALSE}
unique_regions <- unique(monthly_complete$Region)

# Initialize a list to store the plots
plots_list <- list()

# Loop through each region and create the facetted plots
for(region in unique_regions) {
  # Filter the data for the current region
  region_data <- filter(monthly_complete, Region == region)
  
  # Create the ggplot object
  p <- ggplot(region_data, aes(x = Month, y = Percentage, fill = Sex_Species1)) +
    geom_col(position = position_dodge(width = 0.8), width = 0.7) +
    facet_wrap(~Year, ncol = 1, scales = "free_y") +
    scale_fill_brewer(palette = "Paired") +
    labs(title = paste("Monthly Percentage of BHS Detections by Gender in", region, "by Year"),
         x = "Month", 
         y = "Percentage (%)") +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
  
  # Convert to a plotly interactive object
   # Convert to a plotly interactive object
  p_interactive <- ggplotly(p)
  
  # Store the plotly object in the list
  plots_list[[region]] <- p_interactive
  
}

htmltools::tagList(plots_list)
```

*** 

## Percentage of Bighorn Sheep Detections by Age:

The following graph demonstrates the overall detection of Bighorn Sheep by age.
```{r, echo=FALSE}

unique_age_names <- unique(sheep_standardize$Age_Species1)

unique_age_names_no_na <- unique_age_names[!is.na(unique_age_names)]

sheep_standardize$Age_Species1 <- ifelse(sheep_standardize$Age_Species1 %in% c(unique_age_names_no_na), sheep_standardize$Age_Species1, 'Unknown/Other')

coul <- brewer.pal(length(unique_age_names), "Set3")
counts2 <- table(sheep_standardize$Age_Species1)
midpoints <- barplot(counts2, main="Total Count of Bighorn Sheep Detection by Age",
   xlab="Age",
   ylab= 'Total Count',
   col=coul,
   ylim=c(0, max(counts2) + 20)
   )
text(midpoints, counts2, labels=counts2, pos=3, cex=0.8)

```

*** 

## Number of Cameras:

The following represents the park wide camera counts by different years and regions that were located in the provided files.

### Annual Parkwide Number of Cameras:

```{r, include=FALSE}
photoData$Year <- as.numeric(format(as.POSIXct(photoData$DateTime, tz = "UTC"), "%Y"))
```

```{r, include = FALSE}
df2023 <- photoData[photoData$Year == "2023",]
df2022 <- photoData[photoData$Year == "2022",]
df2021 <- photoData[photoData$Year == "2021",]
```


```{r, include=FALSE}
cameras_2023 <- sum(df2023$Camera)
cameras_2023
```


```{r, include=FALSE}
cameras_2022 <- sum(df2022$Camera)
cameras_2022
```

```{r, include=FALSE}
cameras_2021 <- sum(df2021$Camera)
cameras_2021
```


```{r, echo=FALSE}
labels <- c("2021", "2022", "2023")
camera_data <- data.frame(camera_sum = c(cameras_2021, cameras_2022, cameras_2023), Label = labels)
aggregate_data <- aggregate(camera_sum ~ Label, data = camera_data, FUN = sum)
bar_colors <- c("pink", "lightgreen", "skyblue")

num_cameras <- barplot(aggregate_data$camera_sum, names.arg = aggregate_data$Label, main="Number of Cameras by Year at Capitol Reef",
   xlab="Year",
   ylab= 'Total Count',
   col = bar_colors,
   )
text( x = 1:length(aggregate_data$camera_sum), y = aggregate_data$camera_sum,
     label = aggregate_data$camera_sum, pos = 3, cex = 0.8, col = "black")
```

### Number of cameras per month by year overall 

```{r, include=FALSE}
photoData$month <- as.numeric(format(as.POSIXct(photoData$DateTime, tz = "UTC"), "%m"))
```
```{r, include=FALSE}
library(dplyr)

grouped_data <- photoData %>%
  group_by(Year, month) %>%
  summarise(count = n())
```
```{r, include=FALSE}
grouped_data$month <- factor(grouped_data$month, levels = 1:12, labels = month.name)
```
```{r, echo=FALSE}
library(ggplot2)

ggplot(grouped_data, aes(x = month, y = count, fill = factor(Year))) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(x = "Month", y = "Total Count", fill = "Year") +
  geom_text(aes(label = count), size = 2, position = position_dodge(width = 0.9), vjust = -0.5) +
  ggtitle("Number of Cameras per Month by Year") +
  theme_minimal() +
  theme(legend.position = "top", axis.text.x = element_text(angle = 45, hjust = 1))

```


*** 

## Percentage of Detection By Species

The following figures represent the percentage of detection by species between `r format(min_date, '%m/%d/%Y')` and `r format(max_date, '%m/%d/%Y')`.

```{r photoData_preprocessing, include=FALSE}

# Create a new column called "month" extracting month from DateTime
photoData$month <- month(photoData$DateTime)

# If you want the month name instead of the month number, you can use month.abb or month.name
photoData$monthName <- month.abb[month(photoData$DateTime)]
```

```{r, merge, include=FALSE}
locationColumns <- c("Project", "Region", "RegionSite", "x_utm", "y_utm", "Site")
merged_data <- photoData
#merged_data <- merge(photoData, siteLocations[, locationColumns], by = "RegionSite")
head(merged_data)
```
```{r, include=FALSE}
species1_proportions <- table(photoData$Species1) / nrow(photoData)

#print(species1_proportions)
```

### Parkwide Percentage of detections by species (All Time):

```{r, echo=FALSE}
# Create a pie chart
pie(species1_proportions, labels = NA, main = "Proportions of Species", col = rainbow(length(species1_proportions)))

# Add labels with adjusted size
legend("topright", legend = names(species1_proportions), fill = rainbow(length(species1_proportions)), cex = 0.8, xpd = TRUE)

```

<!-- ### Parkwide Percentage of detections by species (Annual): -->

<!-- ```{r, include=FALSE} -->
<!-- df2023 <- photoData[photoData$Year == "2023", ] -->
<!-- df2022 <- photoData[photoData$Year == "2022", ] -->
<!-- df2021 <- photoData[photoData$Year == "2021", ] -->
<!-- ``` -->


### Parkwide Species Detection Percentage (2023):

```{r, 2023_plot, echo=FALSE}
species_proportions2023 <- table(df2023$Species1) / nrow(photoData)
# Create a pie chart
pie(species_proportions2023, labels = NA, main = "Proportions of Species (2023)", col = rainbow(length(species_proportions2023)))

# Add labels with adjusted size
legend("topright", legend = names(species_proportions2023), fill = rainbow(length(species_proportions2023)), cex = 0.8, xpd = TRUE)
```

### Parkwide 2023 Monthly Species Detection:

``` {r 2023_monthly, echo=FALSE}
for (m in unique(df2023$monthName)) {
  # Subset data for the current month
  data_month <- subset(df2023, monthName == m)
  
  # Calculate species proportions for the current month
  species_proportions <- table(data_month$Species1) / nrow(data_month)
  
  # Create a pie chart
  pie(species_proportions, labels = NA, main = paste("Proportions of Species (", m, " 2023)", sep = ""), col = rainbow(length(species_proportions)))
  
  # Add legend with adjusted size
  legend("topright", legend = names(species_proportions), fill = rainbow(length(species_proportions)), cex = 0.8, xpd = TRUE)
}
```

### Parkwide Species Detection Percentage (2022):

<!-- ```{r, 2022_plot, echo=FALSE} -->
<!-- species_proportions2022 <- table(df2022$Species1) / nrow(photoData) -->
<!-- # Create a pie chart -->
<!-- pie(species_proportions2022, labels = NA, main = "Proportions of Species (2022)", col = rainbow(length(species_proportions2022))) -->

<!-- # Add labels with adjusted size -->
<!-- legend("topright", legend = names(species_proportions2022), fill = rainbow(length(species_proportions2022)), cex = 0.8, xpd = TRUE) -->
<!-- ``` -->

<!-- ### Parkwide 2022 Monthly Species Detection: -->

<!-- ``` {r 2022_monthly, echo=FALSE} -->
<!-- for (m in unique(df2022$monthName)) { -->
<!--   # Subset data for the current month -->
<!--   data_month <- subset(df2022, monthName == m) -->

<!--   # Calculate species proportions for the current month -->
<!--   species_proportions <- table(data_month$Species1) / nrow(data_month) -->

<!--   # Create a pie chart -->
<!--   pie(species_proportions, labels = NA, main = paste("Proportions of Species (", m, " 2022)", sep = ""), col = rainbow(length(species_proportions))) -->

<!--   # Add legend with adjusted size -->
<!--   legend("topright", legend = names(species_proportions), fill = rainbow(length(species_proportions)), cex = 0.8, xpd = TRUE) -->
<!-- } -->
<!-- ``` -->

### Parkwide Species Detection Percentage (2021):

```{r, 2021_plot, echo=FALSE}
species_proportions2021 <- table(df2021$Species1) / nrow(photoData)
# Create a pie chart
pie(species_proportions2021, labels = NA, main = "Proportions of Species (2021)", col = rainbow(length(species_proportions2021)))

# Add labels with adjusted size
legend("topright", legend = names(species_proportions2021), fill = rainbow(length(species_proportions2021)), cex = 0.8, xpd = TRUE)
```

### Parkwide 2021 Monthly Species Detection:

``` {r 2021_monthly, echo=FALSE}
for (m in unique(df2021$monthName)) {
  # Subset data for the current month
  data_month <- subset(df2021, monthName == m)
  
  # Calculate species proportions for the current month
  species_proportions <- table(data_month$Species1) / nrow(data_month)
  
  # Create a pie chart
  pie(species_proportions, labels = NA, main = paste("Proportions of Species (", m, " 2021)", sep = ""), col = rainbow(length(species_proportions)))
  
  # Add legend with adjusted size
  legend("topright", legend = names(species_proportions), fill = rainbow(length(species_proportions)), cex = 0.8, xpd = TRUE)
}
```

### Species Detection by Region — Overall:

``` {r overall_region, echo=FALSE}
for (r in unique(photoData$Region)) {
  # Subset data for the current region
  data_region <- subset(photoData, Region == r)
  
  # Calculate species proportions for the current region
  species_proportions <- table(data_region$Species1) / nrow(data_region)
  
  # Create a pie chart
  pie(species_proportions, labels = NA, main = paste("Proportions of Species ", r , sep = ""), col = rainbow(length(species_proportions)))
  
  # Add legend with adjusted size
  legend("topright", legend = names(species_proportions), fill = rainbow(length(species_proportions)), cex = 0.8, xpd = TRUE)
}
```

